Psychological Needs During Intergroup Contact

### 2021-10-04 PreregistrationAOT-M.Rmd
sampling plan added

@@

@@

@@

**Registration prior to any human observation of the data**. As of the date of submission, the data exist but have not yet been quantified, constructed, observed, or reported by anyone - including individuals that are not associated with the proposed study. Examples include museum specimens that have not been measured and data that have been collected by non-human collectors and are inaccessible.

@@

@@

@@

@@

@@

>>>>

1. \label{hyp:contact}Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes.
    a. \label{hyp:contactFreq}Participants with more intergroup interactions should have a more favorable outgroup attitudes.
    b. \label{hyp:contactDummy}Outgroup attitudes should be higher after an intergroup interaction compared to a time with no outgroup interaction. OR: Outgroup attitudes should increase more after an intergroup interaction compared to a time with no outgroup interaction.
    c. \label{hyp:contactFreqQual}Participants with more intergroup interactions should have a more favorable outgroup attitudes depending on the average interaction quality.
    d. \label{hyp:contactQualCor}Alternative to H\ref{hyp:contactFreqQual}: Interaction quality should relate positively with outgroup attitudes.
2. \label{hyp:AllportsConditions}Based on Allport's optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
    a. \label{hyp:AllportsPred}Based on Allport's optimal contact conditions, outgroup attitudes should be more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
    b. \label{hyp:AllportsQuality}Based on past research on the role of interaction quality, interaction quality should be more perceived as more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
    c. \label{hyp:AllportsQualityMediation}Based on past research on the role of interaction quality, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extend be assumed by interaction quality.
3. \label{hyp:keyNeed}Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
    a. \label{hyp:keyNeedPred}Outgroup attitudes should be more favorable after intergroup interactions with high key need fulfillment.
    b. \label{hyp:keyNeedQual}Interaction Quality should be perceived as more positive after intergroup interactions with higher key need fulfillment.
    c. \label{hyp:keyNeedMediation}The variance explained in outgroup attitudes by key need fulfillment should to a large extend be assumed by interaction quality.
    d. \label{hyp:keyNeedContactType}The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should stronger for intergroup interact than for ingroup interactions.
    e. \label{hyp:keyNeedSDT}The effect of key need fulfillment on outgroup attitudes should be persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
4. \label{hyp:comparison}Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict outgroup attitudes at least as well as Allport's conditions.
    a. \label{hyp:compModel}The need model (H\ref{hyp:keyNeedPred}) should predict more variance in outgroup attitudes than the model based on Allport's conditions (H\ref{hyp:AllportsPred}).
    b. \label{hyp:compTogether}The  effect of key need fulfillment on outgroup attitudes should  persist even when taking other Allport's conditions into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for equal status, common goals, collaboration, and structural support.

@@

This study used an extensive longitudinal design. Using a daily diary format, for at least 30 days participants received a short survey twice per day (at around 12pm and 7pm). We additionally included a longer pre- and post measurement survey the days before and after the extensive longitudinal data collection.

@@

No randomization is involved in this study.

@@

@@

The data was collected as part of a larger collaboration on daily intergroup relations. The researchers have had no direct access to the data, as the data was collected using an automated system using the FormR framework [@Arslan2020].

@@

The data was collected in a three step procedure:

1. Entry Survey: A pre-measurement questionnaire (appr. 25 minutes) including demographic information, and relations to the Dutch majority. (payment: 2 Euros)
2. Experience Recaps: At least 30 days of short reflection surveys (appr. 3—5 minutes) on intergroup interactions twice a day. (payment: 1 Euro per Recap; up to 2 Euros per day)
3. Conclusion Survey: On the last day, we conclude with a post-measurement questionnaire (appr. 25 minutes) with some questions on habits and reflection of the study. (payment: 2 Euros)

After the initial 30 day duration, participants are offered the possibility to continue participating in the study either with payment if daily diary measures were missed during the initial study phase or without payment after a total of 60 daily diary measurements were completed. After the initial 30-day period, participants receive automated feedback based on their own well-being, attitudes, and motive responses as an additional initiative give participants access to their own data and to compensate study participation.

@@

Our target sample size is a sum of 4,000 daily diary measurements. With 100% completion rate that would be archieved with 67 participants (60 daily diary responses each). Given that we expect some incomplete daily diary measurements, we aimed to recruit 80 participants.

@@

The targeted sample size depended on a combination of different factors. Different analyses were planned as part of the collaboration and budgeting was a practical constraint. Some analyses were planned based on (1) the pre- to post measurements, (2) the dynamic developments over the daily diary measurements, as well (3) the contemporary effects within the daily diary measurements.

Power considerations of mixed effects model such as with extensive longitudinal data are difficult to estimate because of the complex covariance structures. Simulation studies based on the first two samples within this project indicated that with well-distributed scales, and small to medium effect sizes, 70-80 participants with at least seven daily diary measurements and a simple pre--post survey were sufficiently powerful (power = .8, alpha = .05) to answer most of the key research questions of the collaboration.

The ultimate sampling procedure decision was made as a practical balancing of the number of participants and the number of measurements provided by each participants. Given that this study primarily focuses on the contemporaneous effects within the daily diary measurements, power should thus be high enough.

@@

Participants will be recruited until 80 participants finish the pre-measurement. Invitations to complete additional daily diary measurements will be extended until participants chose to leave the study or at the most until two months (i.e., 64 days) after the initial entry survey (i.e., from the pre measurement survey).
